---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(dplyr)
library(here)
library(reshape2)
library(stringr)
library(clipr)
#library(gridGraphviz)

`%ni%` = Negate(`%in%`)
```

Protein domain data from IsoSeq
```{r}
domains <- read.csv(here::here("isoseq/data/domains_19012023/domain_analysis.csv"))
domains_gained <- read.csv(here::here("isoseq/data/domains_19012023/domainsgained.csv"))

domains_gained %>% 
  dplyr::group_by(gained, AC) %>% 
  dplyr::summarize(n = n())# %>% View()

domains_gained$ID_version <- sapply(domains_gained$AC, function(x) stringr::str_split(x, "[.]") %>% purrr::map_chr(.,1))


domains_lost <- read.csv(here::here("isoseq/data/domains_19012023/domainslost.csv"))

domains_lost %>% 
  dplyr::group_by(lost, AC) %>% 
  dplyr::summarize(n = n())# %>% View()

domains_lost$ID_version <- sapply(domains_lost$AC, function(x) stringr::str_split(x, "[.]") %>% purrr::map_chr(.,1))
```

gProfiler of genes with domain losses/gains
```{r}
library(gprofiler2)
domains_gained$gene_id <- sapply(domains_gained$gene_id, function(x) str_split(x, "[.]") %>% purrr::map_chr(.,1))

domain_gain_gprof <- list()
for (cond in c("calb", "polyic", "saur", "lps")) {
  go <- gprofiler2::gost(query = domains_gained$gene_id[domains_gained$condition_2 == cond],
                       organism = "hsapiens", ordered_query = FALSE,
                       multi_query = FALSE, significant = T, exclude_iea = FALSE,
                       measure_underrepresentation = FALSE, evcodes = FALSE,
                       user_threshold = 0.05, correction_method = "g_SCS",
                       domain_scope = "annotated", custom_bg = NULL,
                       numeric_ns = "", sources = c("GO:BP", "GO:CC", "GO:MF"), as_short_link = FALSE)
  domain_gain_gprof <- rlist::list.append(
    domain_gain_gprof,
    go$result %>% mutate(condition = cond)
  )
}

venn::venn(sapply(domain_gain_gprof, function(x) return(x$term_name)))

library(VennDiagram)
overlap <- calculate.overlap(sapply(domain_gain_gprof, function(x) return(x$term_name)))

do.call(rbind, domain_gain_gprof) %>% View()

go <- gprofiler2::gost(query = domains_gained$gene_id %>% unique(),
                       organism = "hsapiens", ordered_query = FALSE,
                       multi_query = FALSE, significant = T, exclude_iea = FALSE,
                       measure_underrepresentation = FALSE, evcodes = FALSE,
                       user_threshold = 0.05, correction_method = "g_SCS",
                       domain_scope = "annotated", custom_bg = NULL,
                       numeric_ns = "", sources = c("GO:BP"), as_short_link = FALSE)

go <- gprofiler2::gost(query = domains_lost$gene_id %>% unique(),
                       organism = "hsapiens", ordered_query = FALSE,
                       multi_query = FALSE, significant = T, exclude_iea = FALSE,
                       measure_underrepresentation = FALSE, evcodes = FALSE,
                       user_threshold = 0.05, correction_method = "g_SCS",
                       domain_scope = "annotated", custom_bg = NULL,
                       numeric_ns = "", sources = c("GO:BP"), as_short_link = FALSE)
```

Domain gain and loss counts for immune and non-immune genes
```{r}
domains_gained$gained <- as.factor(domains_gained$gained)
domains_gained %>% 
  group_by(gained) %>% 
  dplyr::summarize(n = n()) %>% 
  ggplot(aes(x = forcats::fct_reorder(gained, -n), y = n)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

imm.ass <- readRDS(here::here("gene_lists/imm.ass.rda"))
domains_gained %>% 
  mutate(set = ifelse(gene_name %in% imm.ass$geneName,
                      "immune",
                      "non_immune")) %>% 
  group_by(gained, set, condition_2) %>% 
  dplyr::summarize(n = n()) %>% 
  ggplot(aes(x = forcats::fct_reorder(gained, -n), y = n, fill = set)) +
  geom_col() +
  facet_grid(condition_2~.) +
  scale_fill_viridis_d() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8),
        text = element_text(size = 16),
        legend.position = "top")
```


Protein domain-GO term association with dcGOR
```{r}
library(devtools)
for(pkg in c("dnet","dcGOR")){
    if(pkg %in% rownames(installed.packages())) remove.packages(pkg)
    install_github(repo=paste("hfang-bristol",pkg,sep="/"))
}
library(dcGOR)
```

```{r}
## 1) prepare input and output files
### input files
anno.file <- "http://dcgor.r-forge.r-project.org/data/Algo/HP_anno.txt"
architecture.file <- "http://dcgor.r-forge.r-project.org/data/Algo/SCOP_architecture.txt"
### output files
output.file <- "sf2HPPA.txt"
RData.HIS.customised <- "sf2HPPA.HIS.RData"
## 2) apply dcGO algorithm to infer domain-centric ontology
res <- dcAlgo(anno.file, architecture.file, output.file, ontology="HPPA", feature.mode="supra", fdr.cutoff=0.05, parallel=FALSE)

res_RData <- dcAlgoPropagate(input.file=output.file, ontology="HPPA", output.file=RData.HIS.customised)

domain <- "PF00096"
dcGOR::dcAlgo()

SCOP.sf <- dcRDataLoader('SCOP.sf')
data <- sample(rowNames(SCOP.sf), 50)
eoutput <- dcEnrichment(data, domain="SCOP.sf", ontology="GOMF")
view(eoutput, top_num=10, sortBy="pvalue", details=TRUE)
visEnrichment(eoutput)

Pfam <- dcRDataLoader('Pfam')
data <- sample(rowNames(Pfam), 100)
eoutput <- dcEnrichment(data, domain="Pfam", ontology="GOMF")

res <- dcGOR::dcEnrichment(data = c("PF00096", "PF00076", "PF08028", "PF00441", "PF00005", "PF05822", "PF14836", "PF02984", "PF00023", "PF00443", "PF00970", "PF00175", "PF13606", "PF00134", "PF00400", "PF00619", "PF00271", "PF02770", "PF00656", "PF03009"), 
                           domain = "Pfam", 
                           ontology = c("GOBP"))
visEnrichment(res)
view(res, top_num=10, sortBy="pvalue", details=TRUE)
```

GO term enrichment of gained domains
```{r}
venn::venn(list(domains_gained$gene_name[domains_gained$condition_2 == "calb"],
           domains_gained$gene_name[domains_gained$condition_2 == "polyic"],
           domains_gained$gene_name[domains_gained$condition_2 == "lps"],
           domains_gained$gene_name[domains_gained$condition_2 == "saur"]))

domain_gain_overlap <- VennDiagram::calculate.overlap(list(
  domains_gained$gene_name[domains_gained$condition_2 == "calb"],
  domains_gained$gene_name[domains_gained$condition_2 == "polyic"],
  domains_gained$gene_name[domains_gained$condition_2 == "lps"],
  domains_gained$gene_name[domains_gained$condition_2 == "saur"]))

res_gained <- dcGOR::dcEnrichment(
  data = sapply(domains_gained$AC, function(x) stringr::str_split(x, "[.]") %>% purrr::map_chr(.,1)), 
                           domain = "Pfam", 
                           sizeRange = c(100, 500),
                           ontology = c("GOMF"))

figure <- visEnrichment(res_gained) 

png(file = here::here("isoseq/figures/domains/domain_gained_go_terms.png"),  
    width = 10, height = 10,
    units = "in", res = 300) 
gridGraphviz::grid.graph(figure, newpage = FALSE, nodesOnTop = TRUE)
dev.off()

res_gained_df <- dcGOR::view(res_gained, details=TRUE, top_num = NA) #%>% clipr::write_clip()
write.csv(res_gained_df, here::here("isoseq/output/res_gained_df.csv"))


res_gained_mf <- dcGOR::dcEnrichment(data = sapply(domains_gained$AC, function(x) stringr::str_split(x, "[.]") %>% purrr::map_chr(.,1)), 
                           domain = "Pfam", 
                           ontology = c("GOMF"))
res_gained_mf_df <- dcGOR::view(res_gained_mf, details=TRUE) #%>% clipr::write_clip()
```


```{r}
sapply(res_gained_df$members, function(x) stringr::str_split(x, ",") %>% purrr::map_chr(.,1))
subset(domains_gained, ID_version %in% sapply(res_gained_df$members, function(x) stringr::str_split(x, ",") %>% purrr::map_chr(.,1))) %>% View
all(sapply(domains_gained$members, function(x) stringr::str_split(x, ",") %>% purrr::map_chr(.,1)) %in% domains_lost$ID_version)
```

GO term enrichment of lost domains
```{r}
res_lost <- dcGOR::dcEnrichment(data = sapply(domains_lost$AC, function(x) stringr::str_split(x, "[.]") %>% purrr::map_chr(.,1)), 
                           domain = "Pfam", 
                           sizeRange = c(100, 500),
                           ontology = c("GOBP"))

figure <- dcGOR::visEnrichment(res_lost) 

png(file = here::here("isoseq/figures/domains/domain_lost_go_terms.png"),  
    width = 10, height = 10,
    units = "in", res = 300) 
gridGraphviz::grid.graph(figure, newpage = FALSE, nodesOnTop = TRUE)
dev.off()


res_lost_df <- dcGOR::view(res_lost, details=TRUE, top_num = NA) #%>% clipr::write_clip()
write.csv(res_lost_df, here::here("isoseq/output/res_lost_df.csv"))

res_lost_mf <- dcGOR::dcEnrichment(data = sapply(domains_lost$AC, function(x) stringr::str_split(x, "[.]") %>% purrr::map_chr(.,1)), 
                           domain = "Pfam", 
                           ontology = c("GOMF"))
res_lost_mf_df <- dcGOR::view(res_lost_mf, details=TRUE) #%>% clipr::write_clip()
```

Checking out what proteins are losing electron transport domains
```{r}
sapply(res_lost_df$members, function(x) stringr::str_split(x, ",") %>% purrr::map_chr(.,1))
subset(domains_lost, ID_version %in% as.vector(unlist(sapply(res_lost_df$members, function(x) stringr::str_split(x, ",")))))
subset(domains_lost, ID_version %in% as.vector(unlist(sapply(res_lost_df$members, function(x) stringr::str_split(x, ",")))))$gene_name
all(sapply(res_lost_df$members, function(x) stringr::str_split(x, ",")) %>% unlist() %>% as.vector() %in% domains_lost$ID_version)
```

Translating between pfam domains and gene ontology with pfam2go (from https://supfam.mrc-lmb.cam.ac.uk/SUPERFAMILY/cgi-bin/dcdownload.cgi)
```{r}
pfam_go <- read.delim(here::here("isoseq/data/supfam/pfam2go_mf.txt"))

unique(domains_gained$AC)
gained <- sapply(unique(domains_gained$AC), function(x) stringr::str_split(x, "[.]") %>% purrr::map_chr(.,1)) %>% unlist() %>% as.vector()
all(gained %in% pfam_go$PFAM)

subset(pfam_go, PFAM %in% gained) %>% subset(Information_content > 3) %>% View()

#gained <- as.data.frame(gained)
subset(pfam_go, PFAM %in% gained) %>% subset(Information_content > 3) %>% group_by(GO_name) %>% dplyr::summarize(n = n(), Information_content, PFAM) %>% View

info_threshold <- 1
domains_gained$go <- NA

for (i in 1:nrow(domains_gained)) {
  if (domains_gained$ID_version[i] %in% pfam_go$PFAM) {
    go_association <- pfam_go$GO_name[pfam_go$PFAM == domains_gained$ID_version[i] & pfam_go$Information_content > info_threshold]
    if (length(go_association) != 0) {
      domains_gained$go[i] <-  toString(as.list(go_association)) 
    }
  }
}

domains_gained$go_top <- NA
for (i in 1:nrow(domains_gained)) {
  if (domains_gained$ID_version[i] %in% pfam_go$PFAM) {
    sub <- subset(pfam_go, PFAM == domains_gained$ID_version[i]) 
    go_association <- sub$GO_name[sub$Information_content == max(sub$Information_content)]
  
    if (length(go_association) != 0) {
      domains_gained$go_top[i] <-  toString(as.list(go_association)) 
    }
  }
}
library(scales)
domains_gained$go_top <- as.factor(domains_gained$go_top)
domains_gained %>% 
  subset(!is.na(go_top)) %>% 
  dplyr::group_by(go_top, condition_2) %>% 
  dplyr::summarize(n = n()) %>% 
  ggplot(aes(y = forcats::fct_reorder(go_top, -n), x = n, fill = condition_2)) +
  geom_col(position = "dodge") +
  scale_y_discrete(labels = label_wrap(40)) 
  #facet_wrap(~condition_2, ncol = 2, scales = "free")
```

```{r}
pfam_go %>% 
  subset(Information_content > 2 & Information_content < 2.2) %>% 
  group_by(PFAM) %>% 
  dplyr::summarize(n = n()) %>% 
  ggplot(aes(x = n)) +
  geom_density()

pfam_go %>% 
  group_by(PFAM) %>% 
  dplyr::summarize(n = n()) %>% 
  subset(n)

pfam_go %>% 
  subset(Information_content > 2 & Information_content < 2.2) %>% 
  group_by(GO_ID) %>% 
  dplyr::summarize(n = n(), GO_name) %>% 
  unique() %>% 
  View()

sdfo <- read.delim(here::here("isoseq/data/supfam/SDFO.both.txt"))

sdfo %>% 
  group_by(SDFO_level) %>% 
  dplyr::summarize(n = n())

informative_go <- subset(sdfo, SDFO_level == "Highly  Informative")$GO_id %>% unique()

pfam_go %>% 
  subset(GO_ID %in% med_informative_go) %>% 
  group_by(GO_ID) %>% 
  dplyr::summarize(n = n(), GO_name) %>% unique() %>% View()

gained <- sapply(unique(domains_gained$AC), function(x) stringr::str_split(x, "[.]") %>% purrr::map_chr(.,1)) %>% unlist() %>% as.vector()

t <- subset(pfam_go, PFAM %in% gained) %>% 
  subset(GO_ID %in% informative_go) %>% 
  group_by(GO_name) %>% 
  dplyr::summarize(n = n()) 
t$GO_name <- as.factor(t$GO_name)
t %>% 
  subset(n > 5) %>% 
  ungroup() %>% 
  ggplot(aes(y = forcats::fct_reorder(GO_name, n), x = n)) +
  geom_col(stat="identity")

```

Normalizing the occurence of domain counts by their GO-domain association number
```{r}
pfam_go_count <- pfam_go %>% 
  group_by(PFAM) %>% 
  dplyr::summarize(n_go = n())

pfam_count <- domains_gained %>% 
  group_by(ID_version) %>% 
  dplyr::summarize(n = n())

colnames(pfam_count)[1] <- "PFAM"

full_join(pfam_go_count, pfam_count, by = "PFAM") %>% 
  subset(!is.na(n)) %>%
  mutate(rel = n / n_go) %>% 
  View()
```

There is also a file from GO themselves, which maps pfam domains to GO terms
```{r}
pfamtogo <- read.delim(here::here("isoseq/data/go/pfam2go.txt"), sep = ";", header = F)
colnames(pfamtogo) <- c("domain", "go")
pfamtogo$pfam <- sapply(pfamtogo$domain, function(x) str_split(x, " ") %>% purrr::map_chr(.,1))
pfamtogo$pfam <- sapply(pfamtogo$pfam, function(x) str_split(x, ":") %>% purrr::map_chr(.,2))

pfamtogo$go_desc <- sapply(pfamtogo$domain, function(x) str_split(x, ">") %>% purrr::map_chr(.,2))
```

```{r}
any(unique(domains_gained$ID_version) %in% pfamtogo$pfam)
length(unique(domains_gained$ID_version)[unique(domains_gained$ID_version) %in% pfamtogo$pfam])
subset(pfamtogo, pfam == "PF00641")

colnames(domains_gained)[colnames(domains_gained) == "ID_version"] <- "pfam"
domains_gained_go <- left_join(domains_gained, pfamtogo, by = "pfam")
```


